---
title: "Example Run"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

TODO unclear if this example will make it into the final release or if would do a more generalized example. 

```{r setup, warning=FALSE, message=FALSE}
library(SEM)

library(dplyr)
library(tidyr)
library(ggplot2)
theme_set(theme_bw())
```


Load the example meteorological data. 

```{r}
metdata <- read.csv(system.file("metdata/example_inputs.csv", package = "SEM"))
metdata$time <- seq(as.POSIXct("2005-01-01 00:00"), as.POSIXct("2005-12-31 23:30"), by = 1800) #TODO should figure out a better way to handel this
```


Quick visualize the meteorological data 

```{r}
# reformat into a long df for easy plotting
metdata %>% 
  gather(variable, value, -time) -> 
  long_metdata


long_metdata %>% 
  ggplot(aes(time, value)) + 
  geom_line() + 
  facet_wrap("variable", scales = "free")
```

# Control Run
Run SEM 

```{r}
pest0 <- c("phloem" = 0, "xylem" = 0, "leaf" = 0, "root" = 0, "stem" = 0)
SEMout_pest0 <- run_SEM(pest = pest0, 
                        pest.time = NULL, 
                        inputs = metdata, 
                        X = pools,              # SEM::pools
                        param_df = params_df)   # SEM::params_df
```


Take a look at SEM results. 

GPP from the entire run. 

```{r}
ggplot(data = SEMout_pest0, aes(time, GPP)) + 
  geom_line() 
```


GPP for a week in the middle of the year.  

```{r}
SEMout_pest0 %>% 
  filter(time > "2005-08-01" & time <"2005-08-07") -> 
  aug_control 
aug_control %>% 
  ggplot(aes(time, GPP)) + 
  geom_line() + 
  labs(tittle = "GPP from 1st Week in August")
```



# Disturbance Run 


```{r}
pest_leaf <- c("phloem" = 0, "xylem" = 0, "leaf" = 0.8, "root" = 0, "stem" = 0)
SEMout_pest_leaf <- run_SEM(pest = pest_leaf, 
                        pest.time = as.POSIXct("2005-08-02 09:30:00 EDT"), 
                        inputs = metdata, 
                        X = pools,              # SEM::pools
                        param_df = params_df)   # SEM::params_df


```



GPP from the entire run. It is hard to see here.... 

```{r}
ggplot(data = SEMout_pest_leaf, aes(time, GPP)) + 
  geom_line() 
```


```{r}
SEMout_pest_leaf %>% 
  filter(time > "2005-08-01" & time <"2005-08-07") -> 
  aug_disturbance 

ggplot(data = aug_disturbance, aes(time, GPP)) + 
  geom_line() + 
  labs(tittle = "GPP from 1st Week in August Distubance Run")
```

Compare GPP between the two runs. 

```{r}
ggplot() + 
  geom_line(data = aug_control, aes(time, GPP, color = "Control Run"),
            linewidth = 1, alpha = 0.8) + 
  geom_line(data = aug_disturbance, aes(time, GPP, color = "Disturbance Run"), 
            linewidth = 1,  alpha = 0.8) 
```

